Predicting the effect of candidate interventions for mitigating disease spread in asymmetric-mixing metapopulations

### 2024-07-25 ms.tex
variables and cpu analysis updates

\vspace{0.5em}
\begin{center}
{[ TABLE \ref{tab:Variables} ABOUT HERE ]} \\
\end{center}

@@

We evaluate hypothetical interventions to limit the spread of
infection in the Reserve by simulating these interventions using the formal
system explained above. Our main outcome variable is the reduction in simulated
COVID-19 incidence among Indigenous peoples compared to the no-intervention
case, defined 
\begin{equation}
  \Delta I = I(\text{no intervention}) - I(\text{intervention } i).
\end{equation}
\noindent
A preliminary step is to find parameterizations that reproduce the observed
outbreaks of B.1.1 and Zeta within the Reserve to obtain a baseline estimate of
infection size $I(\text{no intervention})$. 
To this end we hypothesize that re-infections could have driven the observed
greater spread of the Zeta strain, though it could also be due to the holiday
season around the end of 2020 and beginning of 2021.

Interventions center around reducing the mobility of Indigenous people between
Dourados city and Reserve; the total change in mobility, denoted $\Delta
\gamma$, is the total reduction of the parameter $\gamma_g$ across all 
groups, i.e., 
\begin{equation}
  \Delta \gamma = \sum_g \Delta \gamma_g.
\end{equation}
\noindent
Different interventions allocate $\Delta \gamma$ to different degrees. 
One strategy we test is the \emph{Equal} strategy, where all $\gamma_g$ are
equal. Another named strategy is the \emph{80-20} strategy, where the
reduction in mobility in group $g$ is 80\% of the total, and reduction in
group $g'$ mobility is 20\% of the total, i.e., $\Delta\gamma_g / \Delta\gamma =
0.8$ and $\Delta\gamma_{g'} / \Delta\gamma = 0.2$. To create alternative
strategies, both the total reduction and the ratio of group-level reduction to 
total reduction can be changed. We finally evaluate the hypothesis that
interventions that may be effective for strains with no re-infection (like
B.1.1) will not be equally effective for strains with re-infection (like Zeta),
and vice-versa. This will highlight the importance of accounting for social and
epidemiological context, alike, in pandemic preparedness via this program of
computational evaluation of intervention strategies.

@@

\begin{table}[h]
  \centering
  \caption{\textbf{Model variables, description and values taken.} The
  compartmental model described in Equation~\ref{eq:DiffEqs} includes several
social mixing, behavioral, and epidemiological variables described here.}
  \label{tab:Variables}
  \begin{tabular}{rlp{1.3in}} 
    \toprule
    variable & description & values \\ 
    \midrule 
    $S_g,I_g,R_g$ & Susceptible, infected, recovered compartments by group & $[0.0,1.0]$ \\
    $c_{gg'},c_{g'g}$ & Asymmetric inter-group contact rates & $[0.0,1.0]$ \\
    $\gamma_g$ & Group-level incidence-contact rate with city residents &
      $\{0.0,0.1,\ldots,0.4\}$ \\
    $\rho$ & Global recovery rate & $[0.0,1.0]$ \\
    $\tau$ & Global ransmission rate & $[0.0,1.0]$ \\
    $r$    & Re-susceptible probability, non-zero only for Zeta & $0.0,0.5$ \\
    $g,g'$ & Reference group ($g$) and out-group $g'$ & Terena, Guarani; Bororó, Jaguapiru \\
    \bottomrule
  \end{tabular} 
\end{table}

>>>>

\vspace{0.5em}
\begin{center}
{[ TABLE \ref{tab:ModelVariables} ABOUT HERE ]} \\
\end{center}

@@

We evaluate hypothetical interventions to limit the spread of
infection in the Reserve by simulating these interventions using the formal
system explained above. Our main outcome variable is the reduction in simulated
COVID-19 incidence among Indigenous peoples compared to the no-intervention
case, defined 
\begin{equation}
  \Delta I = I(\text{no intervention}) - I(\text{intervention } i).
  \label{eq:DeltaI}
\end{equation}
\noindent
A preliminary step is to find parameterizations that reproduce the observed
outbreaks of B.1.1 and Zeta within the Reserve to obtain a baseline estimate of
infection size $I(\text{no intervention})$. 
To this end we hypothesize that re-infections could have driven the observed
greater spread of the Zeta strain, though it could also be due to the holiday
season around the end of 2020 and beginning of 2021.

Interventions center around reducing the mobility of Indigenous people between
Dourados city and Reserve; the total change in mobility, denoted $\Delta
\gamma$, is the total reduction of the parameter $\gamma_g$ across all 
groups, i.e., 
\begin{equation}
  \Delta \gamma = \sum_g \Delta \gamma_g.
  \label{eq:DeltaGamma}
\end{equation}
\noindent
Different interventions allocate $\Delta \gamma$ (Equation~\ref{eq:DeltaGamma})
in different ways to hypothetically result in different case reductions, $\Delta
I$ (Equation~\ref{eq:DeltaI}). 
One strategy we test is the \emph{Equal} strategy, where all $\gamma_g$ are
equal. Another named strategy is the \emph{80-20} strategy, where the
reduction in mobility in group $g$ is 80\% of the total, and reduction in
group $g'$ mobility is 20\% of the total, i.e., $\Delta\gamma_g / \Delta\gamma =
0.8$ and $\Delta\gamma_{g'} / \Delta\gamma = 0.2$. To create alternative
strategies, both the total reduction and the ratio of group-level reduction to 
total reduction can be changed. In addition to total case reduction, we 
measure relative group-level case reduction, $\Delta I_g$, 
for each considered strategy to understand whether case reduction is equitable
between villages and ethnicities under different strategies. 
We finally evaluate the hypothesis that interventions that may be effective for
strains with no re-infection (like B.1.1) will not be equally effective for
strains with re-infection (like Zeta), and vice-versa. 
% This will highlight the importance of accounting for social and
% epidemiological context, alike, in predicting the effectiveness of mitigation
% strategies for pandemic preparedness.

@@

\begin{table}[h]
  \centering
  \caption{\textbf{Model variables, description and values taken.} The
  compartmental model described in Equation~\ref{eq:DiffEqs} includes several
social mixing, behavioral, and epidemiological variables described here.}
  \label{tab:ModelVariables}
  \begin{tabular}{rlp{1.3in}} 
    \toprule
    variable & description & values \\ 
    \midrule 
    $S_g,I_g,R_g$ & Susceptible, infected, recovered compartments by group & $[0.0,1.0]$ \\
    $c_{gg'},c_{g'g}$ & Asymmetric inter-group contact rates & $[0.0,1.0]$ \\
    $\gamma_g$ & Group-level incidence-contact rate with city residents &
      $\{0.0,0.1,\ldots,0.4\}$ \\
    $\rho$ & Global recovery rate & $[0.0,1.0]$ \\
    $\tau$ & Global ransmission rate & $[0.0,1.0]$ \\
    $r$    & Re-susceptible probability, non-zero only for Zeta & $0.0,0.5$ \\
    $g,g'$ & Reference group ($g$) and out-group $g'$ & Terena, Guarani; Bororó, Jaguapiru \\
    \bottomrule
  \end{tabular} 
\end{table}
